Propositions as Sessions, Mechanically

### 2015-04-10 paper/background.tex
Minor fixes. Resolved a todo regarding Lolliproc.

Concurrency is provided by control operators which, respectively, spawn a
child process and wait for a child process to complete. The spawning process
creates a channel between parent and child process; a channel is represented
as a continuation and the type denotes the communication protocol
(i.e. session type) between parent and child \todo{what part of session type
  is being discussed here? spawning process or spawning operation?}.

>>>>

Concurrency is provided by control operators which, respectively, spawn a
child process and wait for a child process to complete. The spawning operator
creates a channel between the parent and child processes; a channel is
represented by a pair of endpoints with dual types denoting the communication
protocol (i.e. session type) between parent and child.

### 2015-04-10 paper/formalisation.tex
Minor fixes. Resolved a todo regarding Lolliproc.

\textbf{Weakening and Contraction} Initially, an implicit notion of weakening
was chosen so as to be in line with \citeauthor{Wadler:2014}'s
definitions. However, explicit weakening is required to differentiate the
principal cut elimination rules for the interactions between servers and
clients; spawning and garbage collection (weakening). The paper presentation
does not specify where weakening could occur; it is tacitly assumed weakening
can always be pushed further up the derivation tree to a base rule
e.g. \coqe$cp_fwd$ rule.

@@

\textbf{Empty Choice} The empty choice rule may only have a single binding in
the context (cf. an arbitrary context in \citeauthor{Wadler:2014}'s
presentation~\cite{Wadler:2014}). Adding an arbitrary environment to the empty
choice constructor would prohibit \coqe$in_env_fv$ since its process no longer
contains all the names of its environment. Not only would weakening be
inadmissible if \coqe$in_env_fv$ did not hold, but also axiom cut reduction:
\begin{coq}
Lemma reduce_axcut:
  forall P A (w : atom) $\Gamma$
         (NFV: w $\notin$ fv_proc P)
         (WT: $\nu$ A.(w $\link$ 0 $\mid$ P) $\tpcp$ $\Gamma$),
    (open_proc P w) $\tpcp$ $\Gamma$.
\end{coq}

@@

An alternative formulation, which avoids the need to handle error cases, is to
first define only the positive cases using an inductive type, the so-called
\textit{graph relation}, and then define a dependent function which only
considers the positive cases within the relation~\cite{McKinna:2009}. However,
several issues were encountered when attempting to specify the
translation. The graph relation requires much of the clutter present in both
the \coqe!cp_rule! and \coqe!wt_tm! definition such as linear enviornments and
cofinite quanitifcation.

i'm not sure about that but what is can i say so for example lets see see no i
also have to thing about the general guidelines i can be giving for this
section so we have something to say about guidelines do we yes of course we
have can say that out of this we realised that it is important for a
representation to be good enough to succintly express that which we wish to
say in particular the graph relation the inductive type that we define
propositionally is better one would think than the other way the monadic
possible failure way is not producing a valid thing that we can inductively
induct on so we concern ourself with the proof of translation preserving the
typed term and i think it is important to remember the problem we had with the
inductive hypothesis we had something do with translating the inner term of a
weakening term in a larger context than the one we had available and that was
to do with the destruction of the term inference thing i think no no it was
something else we had the inner term and we had it well typed in an
environment that was smaller sure and then our thing our function was saying
we had some environment that typed both the inner term and the outer weakening
term but we know that is wrong so wrong so here is perhaps where the inductive
type graph relation thing wins out in that we can have a hypothesis for the
inner term which is sufficiently small for the environment and we pattern
match not only on the term but also on the environment yes that is we will do
because it seems to work only when we get to the messy two the messy two
constructor caases will this become a bigg problem inthat we will then have
all this stuff about environment being split and needing to be recombined in
the composite process that is a term not a process but the point still stands
it will need to be about the term and the splitting is really a duplication of
the unfortunately the well typed term relation but otherwise i cannot think
how to reuse or obtain that information for the translation to cp

Roughly, the theorm aiming to prove is:

>>>>

\textbf{Weakening and Contraction.} Initially, an implicit notion of weakening
was chosen so as to be in line with \citeauthor{Wadler:2014}'s
definitions. However, explicit weakening is required to differentiate the
principal cut elimination rules for the interactions between servers and
clients; spawning and garbage collection (weakening). The paper presentation
does not specify where weakening could occur; it is tacitly assumed weakening
can always be pushed further up the derivation tree to a base rule
e.g. \coqe$cp_fwd$ rule.

@@

\textbf{Empty Choice.} The empty choice rule may only have a single binding in
the context (cf. an arbitrary context in \citeauthor{Wadler:2014}'s
presentation~\cite{Wadler:2014}). Adding an arbitrary environment to the empty
choice constructor would prohibit \coqe$in_env_fv$ since its process no longer
contains all the names of its environment. Not only would weakening be
inadmissible if \coqe$in_env_fv$ did not hold, but also axiom cut reduction:
\begin{coq}
Lemma reduce_axcut:
  forall P A (w : atom) $\Gamma$
         (NFV: w $\notin$ fv_proc P)
         (WT: $\nu$ A.(w $\link$ 0 $\mid$ P) $\tpcp$ $\Gamma$),
    (open_proc P w) $\tpcp$ $\Gamma$.
\end{coq}

@@

An alternative formulation, which avoids the need to handle error cases, is to
first define only the positive cases using an inductive type, the so-called
\textit{graph relation}, and then define a dependent function which only
considers the positive cases within the relation~\cite{McKinna:2009}. However,
several issues were encountered when attempting to specify the
translation. The graph relation requires much of the clutter present in both
the \coqe!cp_rule! and \coqe!wt_tm! definition such as linear enviornments and
cofinite quanitifcation.
\begin{comment}
i'm not sure about that but what is can i say so for example lets see see no i
also have to thing about the general guidelines i can be giving for this
section so we have something to say about guidelines do we yes of course we
have can say that out of this we realised that it is important for a
representation to be good enough to succintly express that which we wish to
say in particular the graph relation the inductive type that we define
propositionally is better one would think than the other way the monadic
possible failure way is not producing a valid thing that we can inductively
induct on so we concern ourself with the proof of translation preserving the
typed term and i think it is important to remember the problem we had with the
inductive hypothesis we had something do with translating the inner term of a
weakening term in a larger context than the one we had available and that was
to do with the destruction of the term inference thing i think no no it was
something else we had the inner term and we had it well typed in an
environment that was smaller sure and then our thing our function was saying
we had some environment that typed both the inner term and the outer weakening
term but we know that is wrong so wrong so here is perhaps where the inductive
type graph relation thing wins out in that we can have a hypothesis for the
inner term which is sufficiently small for the environment and we pattern
match not only on the term but also on the environment yes that is we will do
because it seems to work only when we get to the messy two the messy two
constructor caases will this become a bigg problem inthat we will then have
all this stuff about environment being split and needing to be recombined in
the composite process that is a term not a process but the point still stands
it will need to be about the term and the splitting is really a duplication of
the unfortunately the well typed term relation but otherwise i cannot think
how to reuse or obtain that information for the translation to cp
\end{comment}
Roughly, the theorm aiming to prove is:

### 2015-04-10 paper/introduction.tex
Minor fixes. Resolved a todo regarding Lolliproc.

This project describes work towards the formalisation of calculi developed by
\citeauthor{Wadler:2014}~\cite{Wadler:2014} describing a relationship between
classical linear logic (CP) and a functional language with session types (GV)
using interactive theorem proving (in particular, the system Coq). The
formalisation is a contribution to the practice of programming language
metatheory and informs what is required of a framework for formalising
$\lambda$-calculus type systems with session types. Such a framework should
enable a whole class of related languages and type systems to be formalised
with little extra proof effort and is inspired by previous work on a
mechanised framework for $\pi$-calculus type systems~\cite{Gay:2001:FFP}. The
current work focuses on the $\lambda$-calculus to provide a basis on which
session types for mainstream programming languages can be studied, taking
advantage of previous proof effort through the re-use of existing libraries
for programming language metatheory. The work has presented a number of
interesting challenges for the formalisation of programming languages and as a
result generated a number of useful guidelines and issues to consider.

>>>>

This project describes work towards the formalisation of calculi developed by
\citeauthor{Wadler:2014}~\cite{Wadler:2014} describing a relationship between
classical linear logic (CP) and a functional language with session types (GV)
using interactive theorem proving (in particular, the system Coq). The
formalisation is a contribution to the practice of programming language
metatheory and informs what is required of a framework for formalising
$\lambda$-calculus type systems with session types. Such a framework should
enable a whole class of related languages and type systems to be formalised
with little extra proof effort and is inspired by previous work on a
mechanised framework for $\pi$-calculus type systems~\cite{Gay:2001:FFP}. The
current work focuses on the $\lambda$-calculus to provide a basis on which
session types for mainstream programming languages can be studied, taking
advantage of previous proof effort through the reuse of existing libraries
for programming language metatheory. The work has presented a number of
interesting challenges for the formalisation of programming languages and as a
result generated a number of useful guidelines and issues to consider.
